Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-18 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

A conventional quantum probability measure can be easily constructed
from a state according to Born rule~\citep{Born1983,peres1995quantum,544199,Jaeger2007}.
This state is also the unique state consistent with the constructed
probability measure according to Gleason's theorem~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}:

\begin{thm}[$\mathscr{I}_{\infty}$ Variant of Gleason]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
the state consistent with it on every projector, $\coreBorn\left(\bar{\mu},\events\right)$,
is unique. \end{thm}

\noindent Since QIVPMs are just finite precision generalization of
conventional quantum probability measures, we also want to seek a
simple rule to construct them, and understand which states are consistent
with the constructed QIVPM. However, there exists a QIVPM such that
no state can be consistent with it. Since no state is consistent with
it, there is no natural way to construct it from any state.

@@

In other words, if the measurement intervals are too imprecise and
the QIVPM are defined so wildly, the results we get might contradict
to each other, and exclude any possible state from being consistent
with the measurement results. To address this problem, we are going
to consider a simple rule which can induce well-behaved QIVPMs by
quantum probability measures systematically. Recall when we generalized
quantum probability measures to QIVPMs, we changed the range from
the unit interval~$\left[0,1\right]$ to a set of intervals~$\mathscr{I}$.
This replacement suggests to find a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
to replace a real-valued probability~$\mu\left(P\right)$ by an interval-valued
one~$\iota\left(\mu\left(P\right)\right)$. In other words, compositing
$\iota$ with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
produces an interval-valued function~$\iota\circ\mu:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$
which has the same domain and range as a QIVPM. In order to guarantee
$\iota\circ\mu$ always a QIVPM, we consider the following homomorphism-like
properties for $\iota$.

@@

\noindent Because the similarity between Def.~\ref{def:THOS} and
Def.~\ref{def:QIVPM}, $\iota\circ\mu$ should always be a IVPM no
matter in classical or quantum case. Surprisingly, if $\iota\circ\mu$
is always a IVPM, then $\iota$ must be a THOS rule.

\begin{thm}[The equivalence among THOS rules and IVPMs]\label{thm:iota-statements}The
following three statements are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}The function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is a THOS rule.
\item \label{enu:iota-mu-CIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a classical IVPM for any classical probability measure~$\mu:\events_{C}\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a QIVPM for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
on any Hilbert space~$\Hilb$.
\end{enumerate}
\end{thm}

As we have discussed, Statement~\ref{enu:iota-subject-to} implies
\ref{enu:iota-mu-QIVPM}. The rest of the proof is to prove that if
$\iota$ is not a THOS rule, then $\iota\circ\mu$ might not be a
QIVPM for any quantum probability measure~$\mu$. If $\iota$ doesn't
satisfy Eq.~(\ref{eq:iota-Inclusion}), we can find three particular
numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$ such
that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, $\iota$ is a THOS rule if and only
if $\iota\circ\mu$ is always a QIVPM. Similarly, Statement~\ref{enu:iota-subject-to}
and \ref{enu:iota-mu-CIVPM} are equivalent as well~\citep{HOSTunpublished}.

@@

Since THOS rules have many desirable properties, the intervals it
mapping to either is trivial or contains infinity intervals.

\begin{thm}[{The range of THOS rule}s]\label{thm:convex-uncountable}For
any function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$ described
in Thm.~\ref{thm:iota-statements}, either $\mathscr{I}=\mathscr{I}_{0}$
defined in Eq.(\ref{eq:3-value-intervals}) or $\mathscr{I}$ contains
uncountably many intervals.\end{thm}

>>>>

A conventional quantum probability measure can be easily constructed
from a state according to Born rule~\citep{Born1983,peres1995quantum,544199,Jaeger2007}.
This state is also the unique state consistent with the constructed
probability measure according to Gleason's theorem~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}
which can be reformulated in QIVPM using infinitely precise intervals~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$:

\begin{thm}[$\mathscr{I}_{\infty}$ Variant of Gleason]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}$,
the state consistent with it on every projector, $\coreBorn\left(\bar{\mu},\events\right)$,
is unique. \end{thm}

\noindent When generalizing to finite precision QIVPMs, we also want
to seek a simple rule to construct them, and understand which states
are consistent with them. However, if the measurement intervals are
too imprecise and the QIVPM are defined so wildly, the results we
get might contradict to each other, and exclude any possible state
from being consistent with the measurement results.

@@

To address this problem, we are going to consider a simple rule which
can induce well-behaved QIVPMs by quantum probability measures systematically.
When we generalized quantum probability measures to QIVPMs, we changed
the range from the unit interval~$\left[0,1\right]$ to a set of
intervals~$\mathscr{I}$. This replacement suggests to find a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
to replace a real-valued probability~$\mu\left(P\right)$ by an interval-valued
one~$\iota\left(\mu\left(P\right)\right)$. In other words, compositing
$\iota$ with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
produces a function~$\iota\circ\mu:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$
which has the same domain and range as a QIVPM. In order to guarantee
$\iota\circ\mu$ always a QIVPM, we consider the following homomorphism-like
properties for $\iota$.

@@

\noindent Because the similarity between Def.~\ref{def:THOS} and
Def.~\ref{def:QIVPM}, $\iota\circ\mu$ should always be a IVPM no
matter in classical or quantum case. Conversely, if $\iota\circ\mu$
is always a IVPM, then $\iota$ must be a THOS rule.

\begin{thm}[The equivalence among THOS rules and IVPMs]\label{thm:iota-statements}The
following three statements are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}The function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is a THOS rule.
\item \label{enu:iota-mu-CIVPM}The composited function~$\iota\circ\mu:\events_{C}\rightarrow\mathscr{I}$
is always a classical IVPM for any classical probability measure~$\mu:\events_{C}\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a QIVPM for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
on any Hilbert space~$\Hilb$.
\end{enumerate}
\end{thm}

As we have discussed, Statement~\ref{enu:iota-subject-to} implies
\ref{enu:iota-mu-QIVPM}. The rest of the proof is to prove that if
$\iota$ is not a THOS rule, then $\iota\circ\mu$ might not be a
QIVPM. If $\iota$ doesn't satisfy Eq.~(\ref{eq:iota-Inclusion}),
we can find three particular numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, $\iota$ is a THOS rule if and only
if $\iota\circ\mu$ is always a QIVPM. Similarly, Statement~\ref{enu:iota-subject-to}
and \ref{enu:iota-mu-CIVPM} are equivalent as well~\citep{HOSTunpublished}.

@@

Although THOS rules have many desirable properties, the intervals
it mapping to either is trivial or contains infinity intervals.

\begin{thm}[The range of THOS rules]\label{thm:convex-uncountable}For
any THOS rule~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$, either
$\mathscr{I}=\mathscr{I}_{0}$ defined in Eq.(\ref{eq:3-value-intervals})
or $\mathscr{I}$ contains uncountably many intervals.\end{thm}
